---
title: "Determining descrimination of VAR features"
author: "Gerry Tonkin-Hill"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_width: 12
    fig_height: 8
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=TRUE, warning=FALSE, message=FALSE,
                      tidy=TRUE)
```

###Load Libraries
```{r}
#libraries used for analysis
library(edgeR)
library(limma)
library(quadprog)
library(ruv)

#libraries required for graphics
library(ggplot2)
library(reshape2)
library(EDASeq)
library(RColorBrewer)
library(VennDiagram)
library(data.table)
library(dplyr)
library(stringr)
library(knitr)
library(pheatmap)
library(glmnet)


wd = "./"
colors <- c("#0571b0","#ca0020")
```

To investigate the utility of using the identified features to distinguish severe and non-severe cases we took the up-regulated features in each category (transcript, domain or segment) and performed logistic regression with lasso regularization. We used cross-validation to determine the optimal $\lamda$ value and give an indication of the best case discrimination between the cases. It should be noted that the classification accuracy reported here may not generalise to other samples. Due to the lack of additional samples a more extensive validation was not possible.

##Transcript level

###Combined Assembly Transcripts

```{r}
# knit('../transcript_level/TextS2_combined_assembly_transcript_analysis.Rmd',
#      tangle=TRUE,
#      output = './TextS2_combined_assembly_transcript_analysis.R')
setwd("../transcript_level/")
pdf(file = NULL)
source('../prediction_analysis_for_reviewer/TextS2_combined_assembly_transcript_analysis.R')
dev.off()
setwd("../prediction_analysis_for_reviewer/")

select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)
nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resUp),]

x <- t(log2.norm.counts)
y <- substring(colnames(log2.norm.counts),1,1)
y[y=="I"] <- "non-severe"
y[y=="S"] <- "severe"
y <- as.factor(y)
cvfit = cv.glmnet(x, y, family = "binomial", type.measure = "class")
plot(cvfit)
cvfit$lambda.1se
coef(cvfit, s = "lambda.1se")
cvfit$lambda.min
coef(cvfit, s = "lambda.min")
cvfit$cvm[cvfit$lambda==cvfit$lambda.min]
```

###Transcript Clusters
```{r}
rm(list=ls(all=TRUE))

# knit('../transcript_level/TextS3_seperate_assembly_CORSET_analysis.Rmd',
#      tangle=TRUE,
#      output = './TextS3_seperate_assembly_CORSET_analysis.R')
setwd("../transcript_level/")
pdf(file = NULL)
source('../prediction_analysis_for_reviewer/TextS3_seperate_assembly_CORSET_analysis.R')
dev.off()
setwd("../prediction_analysis_for_reviewer/")

select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)
nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(SigUp),]

x <- t(log2.norm.counts)
y <- substring(colnames(log2.norm.counts),1,1)
y[y=="I"] <- "non-severe"
y[y=="S"] <- "severe"
y <- as.factor(y)
cvfit = cv.glmnet(x, y, family = "binomial", type.measure = "class")
plot(cvfit)
cvfit$lambda.1se
coef(cvfit, s = "lambda.1se")
cvfit$lambda.min
coef(cvfit, s = "lambda.min")
cvfit$cvm[cvfit$lambda==cvfit$lambda.min]
```

##Domain Level
```{r}
rm(list=ls(all=TRUE))

# knit('../domain_level/TextS4_domain_tree_analysis.Rmd',
#      tangle=TRUE,
#      output = './TextS4_domain_tree_analysis.R')
setwd("../domain_level/")
pdf(file = NULL)
source('../prediction_analysis_for_reviewer/TextS4_domain_tree_analysis.R')
dev.off()
setwd("../prediction_analysis_for_reviewer/")

select <- order(rowMeans(counts(dds_dom, normalized=TRUE)), decreasing=TRUE)
nt <- normTransform(dds_dom)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resSigLFC),]

x <- t(log2.norm.counts)
y <- substring(colnames(log2.norm.counts),1,1)
y[y=="I"] <- "non-severe"
y[y=="S"] <- "severe"
y <- as.factor(y)
cvfit = cv.glmnet(x, y, family = "binomial", type.measure = "class")
plot(cvfit)
cvfit$lambda.1se
coef(cvfit, s = "lambda.1se")
cvfit$lambda.min
coef(cvfit, s = "lambda.min")
cvfit$cvm[cvfit$lambda==cvfit$lambda.min]



select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)
nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resUp),]

x2 <- t(log2.norm.counts)
y2 <- substring(colnames(log2.norm.counts),1,1)
y2[y2=="I"] <- "non-severe"
y2[y2=="S"] <- "severe"
y2 <- as.factor(y2)
cvfit2 = cv.glmnet(x2, y2, family = "binomial", type.measure = "class")
plot(cvfit2)
cvfit2$lambda.1se
coef(cvfit2, s = "lambda.1se")
cvfit2$lambda.min
coef(cvfit2, s = "lambda.min")
cvfit2$cvm[cvfit2$lambda==cvfit2$lambda.min]
```

##Segment Level

```{r}
rm(list=ls(all=TRUE))

# knit('../segment_level/TextS6_segment_level_analysis.Rmd',
#      tangle=TRUE,
#      output = './TextS6_segment_level_analysis.R')
setwd("../segment_level/")
pdf(file = NULL)
source('../prediction_analysis_for_reviewer/TextS6_segment_level_analysis.R')
dev.off()
setwd("../prediction_analysis_for_reviewer/")

select <- order(rowMeans(counts(dds_block, normalized=TRUE)), decreasing=TRUE)
nt <- normTransform(dds_block)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resSigLFC),]

x <- t(log2.norm.counts)
y <- substring(colnames(log2.norm.counts),1,1)
y[y=="I"] <- "non-severe"
y[y=="S"] <- "severe"
y <- as.factor(y)
cvfit = cv.glmnet(x, y, family = "binomial", type.measure = "class")
plot(cvfit)
cvfit$lambda.1se
coef(cvfit, s = "lambda.1se")
cvfit$lambda.min
coef(cvfit, s = "lambda.min")
cvfit$cvm[cvfit$lambda==cvfit$lambda.min]


select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)
nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resUp),]

x <- t(log2.norm.counts)
y <- substring(colnames(log2.norm.counts),1,1)
y[y=="I"] <- "non-severe"
y[y=="S"] <- "severe"
y <- as.factor(y)
cvfit = cv.glmnet(x, y, family = "binomial", type.measure = "class")
plot(cvfit)
cvfit$lambda.1se
coef(cvfit, s = "lambda.1se")
cvfit$lambda.min
coef(cvfit, s = "lambda.min")
cvfit$cvm[cvfit$lambda==cvfit$lambda.min]
```